Computer Simulations of The Dynamics of Asymmetric Dimers in Optical Traps of Varying Polarization

### 2024-06-07 Latex_files/Manuscript Dimers/Manuscript_Optical Communications.tex

Ever since their invention in 1986 by Arthur Ashkin \cite{Ashkin_1986}, Optical Traps or Optical Tweezers (OT) have been used extensively in colloidal physics \cite{Leonardo_2008, Koehler_2011, Curran_1999}. Of particular interest has been the use of a single trapped microsphere as a tracer to probe the rheological properties of the fluid of suspension \cite{Atakhorrami_2006}. However, while working with dense colloidal suspensions, one often ends up trapping more than one microbead in a Gaussian beam OT. Li and Arlt \cite{Li_2008} studied the case of two microspheres trapped in a single OT and opined that multiple trapped beads could be mistaken for a single trapped bead with altered trap stiffness. Theoretical studies on the case of two trapped microspheres by Xu et.al., \cite{Xu_2005} employed a ray-optics based model to show that the two trapped beads are brought into physical contact with each other by optical forces and they also calculated the axial equilibrium positions of the two trapped beads as a function of their size. Sheng-Hua et al \cite{ShengHua_2005} assumed that the two trapped beads experience different trap stiffnesses and went on to show that the trapped beads collide. Experiments in \cite{Praveen_2016} confirmed that the two trapped beads indeed experience different trap stiffnesses when trapped in a single Gaussian beam OT.

Besides multiple trapped beads, trapping and dynamics of asymmetric shapes have generated a lot of interest. Mihiretie et.al., \cite{Loudet_2014} studied the trapping of ellipsoidal microparticles and found that particles of aspect ratio ‘k’ greater than 3 and less than 0.3 exhibited oscillatory dynamics with a coupling between translation and rotation in the trap and those with aspect ratio outside this range exhibited less pronounced oscillations with the major axis pre-dominantly pointed along the laser propagation direction. Cao et.al., \cite{Curran_1999} performed T-Matrix based simulations of the trapping dynamics of nanowires and microcylinders in the Rayleigh regime and found that polarization dictated the final equilibrium orientation with the nanowires equilibrating with a large tilt angle about the laser propagation direction when the beam was linearly polarized and along the beam propagation direction in circularly polarized traps. Chetana D et, al., \cite{Chetana_2022} increased the complexity of the problem by considering an object with both shape as well as optical anisotropy in the form of a chicken Red Blood Cell (cRBC) and showed that the trapping dynamics and the final equilibrium orientation of the cRBC was decided by the beam polarization with the cell showing a dual reorientation behaviour in a linearly polarized trap and a partial second reorientation in a circularly polarized trap.

@@

\paragraph{Simulating motion within an optical trap}
We assume the dimers to be low-Reynold’s number objects and write down their dynamics within the trap using the Langevin equation:
\begin{align}
\label{eq:langevin}
\frac{dx(t)}{dt}=\frac{F(x)}{\gamma}+\sqrt{2D}\eta_x(t)
\end{align}
where $F(x)$ is the optical force directed towards the focal point of the laser However, this is only accurate when close to the equilibrium position, if we wish to probe the phase space we need a means of computing the optical forces for any position or orientation. Fortunately, Vigilante et~al.\ \cite{Vigilante_2020} have already built the ground work for simulating multi-sphere trapping. This work is an extension of the simulation work conducted by \cite{Vigilante_2020} into multi-sphere objects; their paper already has an excellent breakdown of the mathematics that go into calculating the optical and hydrodynamic forces, as such we will only refer to the mathematical methodology when necessary or when our work diverges from \cite{Vigilante_2020}. For simulating asymmetric objects, in which translation and rotation dynamics are coupled, it is useful to have two separate reference systems: The laboratory frame, which refers to the larger system centred at the focal point of the trapping laser; and the particle reference frame which defines the positions of individual spheres, in the case of a dimer the spheres are positioned so that the centre of diffusion is at the origin of the particle reference frame. The axis of the particle reference frame is used to define the orientation of the dimer; with Uz intersecting the centres of the two spheres; any rotations are applied to the reference axis so that the orientation is preserved.

>>>>

Ever since their invention in 1986 by Arthur Ashkin
\cite{Ashkin_1986}, Optical Traps or Optical Tweezers (OT) have been
used extensively in colloidal physics
\cite{Leonardo_2008,Koehler_2011, Curran_1999}.  Of particular
interest has been the use of a single trapped microsphere as a tracer
to probe the rheological properties of the fluid of suspension
\cite{Atakhorrami_2006}.  However, while working with dense colloidal
suspensions, one often ends up trapping more than one microbead in a
Gaussian beam OT.  Li and Arlt \cite{Li_2008} studied the case of two
microspheres trapped in a single OT and opined that multiple trapped
beads could be mistaken for a single trapped bead with altered trap
stiffness. Theoretical studies on the case of two trapped microspheres
by Xu et~al.\ \cite{Xu_2005} employed a ray-optics based model to show
that the two trapped beads are brought into physical contact with each
other by optical forces and they also calculated the axial equilibrium
positions of the two trapped beads as a function of their size.
Sheng-Hua et~al.\ \cite{ShengHua_2005} assumed that the two trapped
beads experience different trap stiffnesses and went on to show that
the trapped beads collide.  Experiments in \cite{Praveen_2016}
confirmed that the two trapped beads indeed experience different trap
stiffnesses when trapped in a single Gaussian beam OT.

Besides multiple trapped beads, trapping and dynamics of asymmetric
shapes have generated a lot of interest.  Mihiretie et~al.\
\cite{Loudet_2014} studied the trapping of ellipsoidal microparticles
and found that particles of aspect ratio ‘k’ greater than $3$ and less
than $0.3$ exhibited oscillatory dynamics with a coupling between
translation and rotation in the trap and those with aspect ratio
outside this range exhibited less pronounced oscillations with the
major axis pre-dominantly pointed along the laser propagation
direction.  Cao et~al.\ \cite{Curran_1999} performed T-Matrix based
simulations of the trapping dynamics of nanowires and microcylinders
in the Rayleigh regime and found that polarization dictated the final
equilibrium orientation with the nanowires equilibrating with a large
tilt angle about the laser propagation direction when the beam was
linearly polarized and along the beam propagation direction in
circularly polarized traps.  Chetana et~al.\ \cite{Chetana_2022}
increased the complexity of the problem by considering an object with
both shape as well as optical anisotropy in the form of a chicken red
bloodc Cell (cRBC) and showed that the trapping dynamics and the final
equilibrium orientation of the cRBC was decided by the beam
polarization with the cell showing a dual reorientation behaviour in a
linearly polarized trap and a partial second reorientation in a
circularly polarized trap.

@@

\paragraph{Simulating motion within an optical trap}
We assume the dimers to be low-Reynolds number objects and write down
their dynamics within the trap using the Langevin equation:
\begin{align}
\label{eq:langevin}
\frac{dx(t)}{dt}=\frac{F(x)}{\gamma}+\sqrt{2D}\eta_x(t)
\end{align}
where $F(x)$ is the optical force directed towards the focal point of the laser However, this is only accurate when close to the equilibrium position, if we wish to probe the phase space we need a means of computing the optical forces for any position or orientation. Fortunately, Vigilante et~al.\ \cite{Vigilante_2020} have already built the ground work for simulating multi-sphere trapping. This work is an extension of the simulation work conducted by \cite{Vigilante_2020} into multi-sphere objects; their paper already has an excellent breakdown of the mathematics that go into calculating the optical and hydrodynamic forces, as such we will only refer to the mathematical methodology when necessary or when our work diverges from \cite{Vigilante_2020}. For simulating asymmetric objects, in which translation and rotation dynamics are coupled, it is useful to have two separate reference systems: The laboratory frame, which refers to the larger system centred at the focal point of the trapping laser; and the particle reference frame which defines the positions of individual spheres, in the case of a dimer the spheres are positioned so that the centre of diffusion is at the origin of the particle reference frame. The axis of the particle reference frame is used to define the orientation of the dimer; with Uz intersecting the centres of the two spheres; any rotations are applied to the reference axis so that the orientation is preserved.
